Synthesizing Symmetric Lenses

### 2019-02-28 papers/SymmetricLenses/main.tex
Finished with section 1

In summary, our contributions are:
\begin{itemize}
%
\item We define {\em simple symmetric lenses}, a subclass of symmetric
lenses that do not maintain any hidden state (\S\ref{sec:overview}).  
%
We present a collection of combinators for building simple symmetric lenses and
describe how to calculate their likelihoods (\S\ref{sec:ssl}).
\item We give an algorithm for synthesizing lenses built from these combinators,
  using a novel application of stochastic regular expressions to guide the
  search process (\S\ref{sec:synthesis}) and prove that our metrics have deep
  semantic meaning.
%
\item We extend the Boomerang implementation with simple symmetric lenses,
  demonstrating that they integrate smoothly with other lens extensions such as
  quotient lenses~\cite{quotientlenses} and matching
  lenses~\cite{matchinglenses}.

\item We also extend Boomerang with symmetric lens synthesis specifications:
  including types, examples, and relevance annotations. We evaluate our implementation, the effects of
  optimizations, and our inclusion of relevance annotations on a set of
  48 lens synthesis benchmarks drawn from data cleaning, view maintenance, and
  file synchronization tasks. The system can synthesize simple symmetric lenses
  for all of the benchmarks in under 30 seconds (\S\ref{sec:evaluation}).
%
\item We show that simple symmetric lenses fall between asymmetric
lenses~\cite{Focal2005-long} and full symmetric
lenses~\cite{symmetric-lenses} in expressiveness.  Indeed, the
class of simple lenses can be characterized semantically as the subset of
full symmetric lenses for which a simple ``forgetfulness'' property holds
(\S\ref{sec:related}).
\end{itemize}
Along the
way, we show the {\em star-semiring equivalences} on regular expressions
remain valid when extended to stochastic regular expressions, preserving
probability distributions as well as languages (\S\ref{sec:sre}).
We close with related work (\S\ref{sec:related}) and
concluding thoughts (\S\ref{sec:conc}).
Elided proofs can be found in the appendices
\ifappendices\else included with the supplemental material\fi.

@@

\begin{theorem}
  If $\Regex \SSREquiv \RegexAlt$ then $\ProbabilityOf{\Regex}{\String} =
  \ProbabilityOf{\RegexAlt}{\String}$, for all strings $\String \in \LanguageOf{\Regex}$.
\end{theorem}
\begin{proof}
  See the appendix %
\ifappendices  Theorem~\ref{thm:semantics-correctness} \else{} included with the supplemental
material\fi.
\end{proof}

This theorem will come in handy as we traverse (with \Expand) or normalize
across (as part of \GreedySynth) star-semiring equivalences.

@@

\subsection{Lens Costs}
\label{subsec:lenscosts}
Our cost metric is based on the expected amount of information required to
recover a string in one data format from the other. We use the function
$\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ to calculate bounds on the
expected amount of information required to recover a string in $\RegexAlt$ from
a string in $\Regex$, synchronized by $\Lens$. Similarly, we use the function
$\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ to calculate bounds on the
expected amount of information required to recover a string in $\Regex$ from a
string in $\RegexAlt$, synchronized by $\Lens$. We write $a(b,c)$ to
mean $(a*b,a*c)$, and $(a,b)+(c,d)$ to mean $(a+c,b+d)$.
\begin{center}
  \begin{tabular}{rcl}
    $\REntropyOf{\RegexAlt \Given \IdentityLensOf{\RegexAlt}, \RegexAlt}$
    & =
    & [0,0]\\

>>>>

In summary, our contributions are as follows:
\begin{itemize}
%
\item We define {\em simple symmetric lenses}, a subclass of symmetric
lenses with no hidden state,
%
we propose a syntax for them---a collection
of combinators for building simple symmetric lenses---and we
describe how to calculate their likelihoods
(\S\ref{sec:overview}--\S\ref{sec:ssl}).  
\item We give an algorithm for synthesizing lenses built from these combinators,
  using a novel application of stochastic regular expressions to guide the
  search process (\S\ref{sec:synthesis}) and prove that our metrics have deep
  semantic meaning. \bcp{``Semantic meaning'' is redundant.  And ``deep
    semantic meaning'' just sounds silly.  Can we say it more precisely?}
%
\item We extend the Boomerang implementation with simple symmetric lenses
and lens synthesis specifications, 
  including types, examples, and relevance annotations. We evaluate our
  implementation, the effects of
  optimizations, and our inclusion of relevance annotations on a set of
  48 lens synthesis benchmarks drawn from data cleaning, view maintenance, and
  file synchronization tasks. The system can synthesize simple symmetric lenses
  for all of the benchmarks in under 30 seconds (\S\ref{sec:evaluation}).
%
\item We prove that simple symmetric lenses fall between asymmetric
lenses~\cite{Focal2005-long} and full symmetric
lenses~\cite{symmetric-lenses} in expressiveness.  Indeed, the
class of simple lenses can be characterized semantically as the subset of
full symmetric lenses for which a simple ``forgetfulness'' property holds
(\S\ref{sec:related}).
\end{itemize}
Along the
way, we show that {\em star-semiring equivalences}\bcp{citation} on regular
expressions 
remain valid when extended to stochastic regular expressions, preserving
probability distributions as well as languages (\S\ref{sec:sre}).
We close with related work (\S\ref{sec:related}) and
concluding thoughts (\S\ref{sec:conc}).
Elided proofs can be found in the appendices
\ifappendices\else included with the supplemental material\fi.

@@

\begin{theorem}
  If $\Regex \SSREquiv \RegexAlt$ then $\ProbabilityOf{\Regex}{\String} =
  \ProbabilityOf{\RegexAlt}{\String}$, for all strings $\String \in \LanguageOf{\Regex}$.
\end{theorem}

(For the proof, see the appendix \ifappendices
Theorem~\ref{thm:semantics-correctness} \else{} included with the
supplemental material\fi.)  This theorem will come in handy as we traverse
(with \Expand) or normalize across (as part of \GreedySynth) star-semiring
equivalences.

@@

\subsection{Lens Costs}
\label{subsec:lenscosts} \bcp{In the intro, we talked about ``likelihood.''
Now we seem to be switching the terminology to ``cost.''}
Our cost metric is based on the expected amount of information required to
recover a string in one data format from the other. We use the function
$\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ to calculate bounds on the
expected amount of information required to recover a string in $\RegexAlt$ from
a string in $\Regex$, synchronized by $\Lens$. Similarly, we use the function
$\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ to calculate bounds on the
expected amount of information required to recover a string in $\Regex$ from a
string in $\RegexAlt$, synchronized by $\Lens$. We write $a(b,c)$ to
mean $(a*b,a*c)$, and $(a,b)+(c,d)$ to mean $(a+c,b+d)$.
\begin{center}
  \begin{tabular}{rcl}
    $\REntropyOf{\RegexAlt \Given \IdentityLensOf{\RegexAlt}, \RegexAlt}$
    & =
    & [0,0]\\
